Illusion Sensitivity and Phenomenological Control

### 2025-05-21 analysis/1_cleaning.qmd
lesgooo @AnafNeves

---
title: "Illusion Sensitivity and Phenomenological Control"
subtitle: "Preprocessing"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: true
---

@@

# path_ig <- read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_illusion.csv")
path_ig <- read.csv("../data/rawdata_illusion.csv")

data_ig <- path_ig |>
  group_by(Illusion_Type) |> 
  mutate(Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
         Task_Difficulty = abs(Illusion_Difference),
         Condition_Illusion = datawizard::categorize(
           Illusion_Strength, split="quantile", n_groups=4,
           labels=c("Congruent - Strong", "Congruent - Mild", "Incongruent - Mild", "Incongruent - Strong")),
         Condition_Difficulty = datawizard::categorize(Task_Difficulty, split="quantile", n_groups=2, labels=c("Hard", "Easy"))) |> 
  ungroup()

```

@@

## Score Computation

@@

data_ppt$IPIP6_Extraversion <- rowMeans(ipip6[grepl("ext", names(ipip6))])
data_ppt$IPIP6_Conscientiousness <- rowMeans(ipip6[grepl("cons", names(ipip6))])
data_ppt$IPIP6_Neuroticism <- rowMeans(ipip6[grepl("neuro", names(ipip6))])
data_ppt$IPIP6_Openness <- rowMeans(ipip6[grepl("open", names(ipip6))])
data_ppt$IPIP6_HonestyHumility <- rowMeans(ipip6[grepl("honhum", names(ipip6))])
data_ppt$IPIP6_Agreeableness <- rowMeans(ipip6[grepl("agr", names(ipip6))])
```

@@

### Experiment Duration

@@

```{r}
data_outliers <- df |>
  mutate(Outlier = ifelse(RT >= 10, TRUE, FALSE)) |>
  group_by(Participant) |>
  mutate(Outlier = ifelse(RT < 0.150 | standardize(RT, robust = TRUE) > 4, TRUE, Outlier)) |>
  ungroup()
```

@@

The removal of individual trials will also be performed (RT < 0.2 s or > 5 s; Thériault et al., 2024) to mitigate the risk of confounding effects driven by extreme speed or accuracy strategies.

@@

```{r}

# outliers$Hello <- data_ppt |>
#   dplyr::filter(pcs_hello != "Hello")
# 
# outliers$Hello <- outliers$Hello$Participant

```

We removed `r length(outliers$Hello)` (`r insight::format_percent(length(outliers$Hello) / nrow(data_ppt))`) participants that did not reported hearing "Hello" at the beggining of the experiment.

```{r}
# remove after more pilots
# outliers$Balls <- data_ppt |>
#   dplyr::filter(pcs_noballs == TRUE) # remove after more pilots
# 
# # outliers$balls <- data_ppt |>
# #   dplyr::filter(pcs_noballs == "TRUE")
# 
# outliers$Balls <- outliers$Balls$Participant
```

We removed `r length(outliers$Balls)` (`r insight::format_percent(length(outliers$Balls) / nrow(data_ppt))`) participants that responded that no balls were presented in the screen.

```{r}

# outliers$Press <- data_ppt |>
#   dplyr::filter(pcs_press < 5)
# 
# outliers$Press <- outliers$Press$Participant
```

We removed `r length(outliers$Press)` (`r insight::format_percent(length(outliers$Press) / nrow(data_ppt))`) participants that responded that no balls were presented in the screen.

```{r}
data_ppt <- filter(data_ppt, !Participant %in% outliers$Hello)
data_ppt <- filter(data_ppt, !Participant %in% outliers$Balls)
data_ppt <- filter(data_ppt, !Participant %in% outliers$Press)

data_ig <- filter(data_ig, !Participant %in% outliers$Hello)
data_ig <- filter(data_ig, !Participant %in% outliers$Balls)
data_ig <- filter(data_ig, !Participant %in% outliers$Press)
```

@@

```{r}
# data_ppt<- filter(data_ppt, !Participant %in% outliers)
# data_ig <- filter(data_ig, Participant %in% data_ppt$Participant)
```

### Age

```{r}
p_age <- estimate_density(data_ppt$Age) |>
  normalize(select = y) |>
  mutate(y = y * 86) |>  # To match the binwidth
  ggplot(aes(x = x)) +
  geom_histogram(data=data_ppt, aes(x = Age, fill=Gender), bins=28) +
  # geom_line(aes(y = y), color = "orange", linewidth=2) +
  geom_vline(xintercept = mean(data_ppt$Age), color = "red", linewidth=1.5) +
  # geom_label(data = data.frame(x = mean(df$Age) * 1.15, y = 0.95 * 75), aes(y = y), color = "red", label = paste0("Mean = ", format_value(mean(df$Age)))) +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )
p_age
```

### Education

@@

### Ethnicity

@@

### Summary

## Save

@@

write.csv(data_ppt, "../data/data_participants.csv", row.names = FALSE)
write.csv(data_ig, "../data/data_IllusionGame.csv", row.names = FALSE)
```

>>>>

---
title: "Illusion Sensitivity and Phenomenological Control"
subtitle: "Preprocessing"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
execute: 
  cache: false
---

@@

# read.csv("https://raw.githubusercontent.com/RealityBending/IllusionGamePhenomenologicalControl/refs/heads/main/data/rawdata_illusion.csv")
data_ig <- read.csv("../data/rawdata_illusion.csv") |>
  mutate(
    Illusion_Effect = ifelse(sign(Illusion_Strength) > 0, "Incongruent", "Congruent"),
    Task_Difficulty = abs(Illusion_Difference),
    Condition_Illusion = case_when(
        Illusion_Effect == "Congruent" ~ "Congruent",
        Illusion_Effect == "Incongruent" & Illusion_Strength <= median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Mild",
        Illusion_Effect == "Incongruent" & Illusion_Strength > median(Illusion_Strength[Illusion_Effect == "Incongruent"], na.rm = TRUE) ~ "Strong"),
    Condition_Difficulty = datawizard::categorize(
      Task_Difficulty, 
      split = "quantile", 
      n_groups = 2, 
      labels = c("Hard", "Easy")),
    .by = "Illusion_Type"
  )
```

@@

## Experiment Duration

## Personality Questionnaires Computation

@@

data_ppt$IPIP6_Extraversion <- rowMeans(ipip6[grepl("Ext", names(ipip6))])
data_ppt$IPIP6_Conscientiousness <- rowMeans(ipip6[grepl("Cons", names(ipip6))])
data_ppt$IPIP6_Neuroticism <- rowMeans(ipip6[grepl("Neuro", names(ipip6))])
data_ppt$IPIP6_Openness <- rowMeans(ipip6[grepl("Open", names(ipip6))])
data_ppt$IPIP6_HonestyHumility <- rowMeans(ipip6[grepl("HonHum", names(ipip6))])
data_ppt$IPIP6_Agreeableness <- rowMeans(ipip6[grepl("Agr", names(ipip6))])
```

@@

@@

@@

We removed individual extreme trials (RT < 0.2 s or > 5 s; Thériault et al., 2024) to mitigate the risk of confounding effects driven by extreme speed or accuracy strategies.

@@

```{r}
outliers_pcs1 <- data_ppt[data_ppt$pcs_hello != "Hello", "Participant"]
outliers_pcs2 <- data_ppt[data_ppt$pcs_noballs == TRUE, "Participant"]
outliers_pcs3 <- data_ppt[data_ppt$pcs_press < 5, "Participant"]
outliers_pcs <- unique(c(outliers_pcs1, outliers_pcs2, outliers_pcs3))
```

We removed `r length(outliers_pcs)` (`r insight::format_percent(length(outliers_pcs) / nrow(data_ppt))`) participants that either did not reported hearing "Hello" at the beginning of the experiment, responded that no balls were presented in the screen, or pressed less than 5 times when asked to do so.

```{r}
# data_ppt <- filter(data_ppt, !Participant %in% outliers_pcs)
# data_ig <- filter(data_ig, !Participant %in% outliers_pcs)
```

@@

## Age

```{r}
p_age <- data_ppt |>  # To match the binwidth
  ggplot(aes(x = data_ppt)) +
  geom_histogram(aes(x = Age, fill=Gender), bins=28) +
  geom_vline(xintercept = mean(data_ppt$Age), color = "red", linewidth=1.5) +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )
p_age
```

## Education

@@

## Ethnicity

@@

# Save

@@

write.csv(data_ppt, "../data/data_participants.csv", row.names = FALSE)
write.csv(data_ig, "../data/data_illusion.csv", row.names = FALSE)
```
